---
title: "Bootstrap versus BHM simulation study"
output:
  pdf_document: default
  html_document: default
date: "2024-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


## Slice sampling MCMC
Code for slice sampling from Neal et al. (2003)
```{r}
# R FUNCTIONS FOR PERFORMING UNIVARIATE SLICE SAMPLING.
#
# Radford M. Neal, 17 March 2008.
#
# Implements, with slight modifications and extensions, the algorithm described
# in Figures 3 and 5 of the following paper:
#
#   Neal, R. M (2003) "Slice sampling" (with discussion), Annals of Statistics,
#      vol. 31, no. 3, pp. 705-767.
#
# See the documentation for the function uni.slice below for how to use it.
# The function uni.slice.test was used to test the uni.slice function.


# GLOBAL VARIABLES FOR RECORDING PERFORMANCE.

uni.slice.calls <- 0	# Number of calls of the slice sampling function
uni.slice.evals <- 0	# Number of density evaluations done in these calls


# UNIVARIATE SLICE SAMPLING WITH STEPPING OUT AND SHRINKAGE.
#
# Performs a slice sampling update from an initial point to a new point that 
# leaves invariant the distribution with the specified log density function.
#
# Arguments:
#
#   x0    Initial point
#   g     Function returning the log of the probability density (plus constant)
#   w     Size of the steps for creating interval (default 1)
#   m     Limit on steps (default infinite)
#   lower Lower bound on support of the distribution (default -Inf)
#   upper Upper bound on support of the distribution (default +Inf)
#   gx0   Value of g(x0), if known (default is not known)
#
# The log density function may return -Inf for points outside the support 
# of the distribution.  If a lower and/or upper bound is specified for the
# support, the log density function will not be called outside such limits.
#
# The value of this function is the new point sampled, with an attribute
# of "log.density" giving the value of the log density function, g, at this
# point.  Depending on the context, this log density might be passed as the 
# gx0 argument of a future call of uni.slice. 
#
# The global variable uni.slice.calls is incremented by one for each call
# of uni.slice.  The global variable uni.slice.evals is incremented by the
# number of calls made to the g function passed.
#
# WARNING:  If you provide a value for g(x0), it must of course be correct!
# In addition to giving wrong answers, wrong values for gx0 may result in
# the uni.slice function going into an infinite loop.

uni.slice <- function (x0, g, w=1, m=Inf, lower=-Inf, upper=+Inf, gx0=NULL)
{
  # Check the validity of the arguments.
  
  if (!is.numeric(x0) || length(x0)!=1
      || !is.function(g) 
      || !is.numeric(w) || length(w)!=1 || w<=0 
      || !is.numeric(m) || !is.infinite(m) && (m<=0 || m>1e9 || floor(m)!=m)
      || !is.numeric(lower) || length(lower)!=1 || x0<lower
      || !is.numeric(upper) || length(upper)!=1 || x0>upper
      || upper<=lower 
      || !is.null(gx0) && (!is.numeric(gx0) || length(gx0)!=1))
  { 
    stop ("Invalid slice sampling argument")
  }
  
  # Keep track of the number of calls made to this function.
  
  uni.slice.calls <<- uni.slice.calls + 1
  
  # Find the log density at the initial point, if not already known.
  
  if (is.null(gx0)) 
  { uni.slice.evals <<- uni.slice.evals + 1
  gx0 <- g(x0)
  }
  
  # Determine the slice level, in log terms.
  
  logy <- gx0 - rexp(1)
  
  # Find the initial interval to sample from.
  
  u <- runif(1,0,w)
  L <- x0 - u
  R <- x0 + (w-u)  # should guarantee that x0 is in [L,R], even with roundoff
  
  # Expand the interval until its ends are outside the slice, or until
  # the limit on steps is reached.
  
  if (is.infinite(m))  # no limit on number of steps
  { 
    repeat
    { if (L<=lower) break
      uni.slice.evals <<- uni.slice.evals + 1
      if (g(L)<=logy) break
      L <- L - w
    }
    
    repeat
    { if (R>=upper) break
      uni.slice.evals <<- uni.slice.evals + 1
      if (g(R)<=logy) break
      R <- R + w
    }
  }
  
  else if (m>1)  # limit on steps, bigger than one
  { 
    J <- floor(runif(1,0,m))
    K <- (m-1) - J
    
    while (J>0)
    { if (L<=lower) break
      uni.slice.evals <<- uni.slice.evals + 1
      if (g(L)<=logy) break
      L <- L - w
      J <- J - 1
    }
    
    while (K>0)
    { if (R>=upper) break
      uni.slice.evals <<- uni.slice.evals + 1
      if (g(R)<=logy) break
      R <- R + w
      K <- K - 1
    }
  }
  
  # Shrink interval to lower and upper bounds.
  
  if (L<lower) 
  { L <- lower
  }
  if (R>upper)
  { R <- upper
  }
  
  # Sample from the interval, shrinking it on each rejection.
  
  repeat
  { 
    x1 <- runif(1,L,R)
    
    uni.slice.evals <<- uni.slice.evals + 1
    gx1 <- g(x1)
    
    if (gx1>=logy) break
    
    if (x1>x0) 
    { R <- x1
    }
    else 
    { L <- x1
    }
  }
  
  # Return the point sampled, with its log density attached as an attribute.
  
  attr(x1,"log.density") <- gx1
  return (x1)
  
}

updates <- function (x0, g, reuse=FALSE)
{ 
  uni.slice.calls <<- 0
  uni.slice.evals <<- 0
  
  s <<- numeric(ss)
  x1 <- x0
  s[1] <<- x0
  last.g <- NULL
  
  for (i in 2:ss)
  { for (j in 1:thin)
  { if (reuse)
  { x1 <- uni.slice (x1, g, w=w, m=m, lower=lower, upper=upper, 
                     gx0=last.g)
  last.g <- attr(x1,"log.density")
  }
    else
    { x1 <- uni.slice (x1, g, w=w, m=m, lower=lower, upper=upper)
    }
  }
    s[i] <<- x1
  }
  return(s)
}


```

## Data
Here is the assumed data for the simulation study.

```{r}
set.seed(123)
#simulation study 
N_mcmc <- 20000
n_tasks <- 3
n_models <- 2

Y_obs <- matrix(rep(0, n_models*n_tasks), ncol = n_tasks)
Y_obs[1,] <- c(100,5000,10000)
Y_obs[2,] <-  c(115,5000,10000)
N_obs <- c(200,10000,20000)
```

## Analytical 95% confidence interval for difference in unweighted mean performance
Here is an analytical 95% confidence interval for the difference in means

```{r}
#assume the true p's are .5,.5,.5 for model 1 and .525,.5,.5 for model 2
#thus, model 2 is strictly better

p1 <- Y_obs[1,]/N_obs
mean(p1)
p2 <- Y_obs[2,]/N_obs
mean(p2)
mean(p1)-mean(p2)

#1: analytical 95% CI
#analytical CI for comparison -- contains 0 (error)
mean(p1)-mean(p2)-2*sqrt(sum(p1*(1-p1)/N_obs)/9+sum(p2*(1-p2)/N_obs)/9)
mean(p1)-mean(p2)+2*sqrt(sum(p1*(1-p1)/N_obs)/9+sum(p2*(1-p2)/N_obs)/9)
#length
4*sqrt(sum(p1*(1-p1)/N_obs)/9+sum(p2*(1-p2)/N_obs)/9)

#2:bootstrap 95% CI
mock_test <- c()
#generate mock responses
for(i in 1:n_tasks){
  mock_test[[i]] <- matrix(rep(0,n_models*N_obs[i]), ncol = n_models)
  for(j in 1:n_models){
    one_ind <- sample(seq(N_obs[i]), size = Y_obs[j,i], replace = FALSE)
    mock_test[[i]][one_ind,j] <- 1
  }
}
```


## Bootstrapped 95% confidence interval for difference in unweighted mean performance
```{r}
mock_test <- c()
#generate mock responses
for(i in 1:n_tasks){
  mock_test[[i]] <- matrix(rep(0,n_models*N_obs[i]), ncol = n_models)
  for(j in 1:n_models){
    one_ind <- sample(seq(N_obs[i]), size = Y_obs[j,i], replace = FALSE)
    mock_test[[i]][one_ind,j] <- 1
  }
}

#bootstrap samples
N_boot <- 5000
p_samps <- rep(0, N_boot)
for(i in 1:N_boot){
  task_1_ind <- sample(seq(N_obs[1]), replace = TRUE)
  task_2_ind <- sample(seq(N_obs[2]), replace = TRUE)
  task_3_ind <- sample(seq(N_obs[3]), replace = TRUE)
  
  p_11 <- sum(mock_test[[1]][task_1_ind,1])/N_obs[1]
  p_12 <- sum(mock_test[[1]][task_1_ind,2])/N_obs[1]
  
  p_21 <- sum(mock_test[[2]][task_2_ind,1])/N_obs[2]
  p_22 <- sum(mock_test[[2]][task_2_ind,2])/N_obs[2]
  
  p_31 <- sum(mock_test[[3]][task_3_ind,1])/N_obs[3]
  p_32 <- sum(mock_test[[3]][task_3_ind,2])/N_obs[3]
  
  p_samps[i] <- (p_11+p_21+p_31)/3-(p_12+p_22+p_32)/3
}

#contains 0 -- error
quantile(p_samps, .025)
quantile(p_samps, .975)

```

## 95% credibility interval for diference in unweighted mean task performance
```{r}
cur_alphas_h <- rep(1, n_models)
cur_betas_h  <-  rep(1, n_models)
cur_thetas_h <- matrix(rep(.5, n_models*n_tasks), nrow = n_models)
alpha_samps <- matrix(rep(0, n_models*N_mcmc), nrow = N_mcmc)
beta_samps <- matrix(rep(0, n_models*N_mcmc), nrow = N_mcmc)
theta_samps <- array(rep(0, n_models*N_mcmc*n_tasks), dim = c(n_models, n_tasks,N_mcmc))

#slice sampling constants
w <- .5
thin <- 1
ss <- 2
lower <- 0
upper <- Inf
m <- Inf

# global index variable for Gibbs sampler
index <- 1

log_conditional_post_a1 <- function(a_inp){
  log_llh_score <- 0
  for(j in 1:n_tasks){
    log_llh_score <- log_llh_score + dbeta(cur_thetas_h[index,j], a_inp, cur_betas_h[index], log=TRUE)
  }
  return(log_llh_score+dnorm(a_inp, mean = 2000, sd = 10,log = TRUE))
}
log_conditional_post_a2 <- function(a_inp){
  log_llh_score <- 0
  for(j in 1:n_tasks){
    log_llh_score <- log_llh_score + dbeta(cur_thetas_h[index,j], a_inp, cur_betas_h[index], log=TRUE)
  }
  return(log_llh_score+dnorm(a_inp, mean = 2100, sd = 10,log = TRUE))
}
log_conditional_post_b1 <- function(b_inp){
  log_llh_score <- 0
  for(j in 1:n_tasks){
    log_llh_score <- log_llh_score + dbeta(cur_thetas_h[index,j], cur_alphas_h[index], b_inp, log=TRUE)
  }
  return(log_llh_score+dnorm(b_inp, mean = 2000, sd = 10,log = TRUE))
}
log_conditional_post_b2 <- function(b_inp){
  log_llh_score <- 0
  for(j in 1:n_tasks){
    log_llh_score <- log_llh_score + dbeta(cur_thetas_h[index,j], cur_alphas_h[index], b_inp, log=TRUE)
  }
  return(log_llh_score+dnorm(b_inp, mean = 1900, sd = 10,log = TRUE))
}

#Gibbs sampling
for(iter in 1:N_mcmc){
  #print(iter)
  #sample through thetas given rest using conjugacy
  for(i in 1:n_models){
    cur_alpha <- cur_alphas_h[i]
    cur_beta <- cur_betas_h[i]
    for(j in 1:n_tasks){
      cur_post_alpha <- Y_obs[i,j] + cur_alpha
      cur_post_beta <- N_obs[j] - Y_obs[i,j] + cur_beta
      cur_thetas_h[i,j] <- rbeta(1, cur_post_alpha, cur_post_beta)
    }
  }
  #reset global index
  index <- 1
  #sample alphas given rest with slice sampling

  new_samp <- updates(cur_alphas_h[1],log_conditional_post_a1)
  cur_alphas_h[1] <- new_samp[2]
  index <- index+1
  
  new_samp <- updates(cur_alphas_h[2],log_conditional_post_a2)
  cur_alphas_h[2] <- new_samp[2]
  index <- index+1

  #reset index
  index <- 1
  #sample betas given rest with slice sampling

  new_samp <- updates(cur_betas_h[1],log_conditional_post_b1)
  cur_betas_h[1] <- new_samp[2]
  index <- index+1
  
  new_samp <- updates(cur_betas_h[2],log_conditional_post_b2)
  cur_betas_h[2] <- new_samp[2]
  index <- index+1
  
  #store samples
  alpha_samps[iter,] <- cur_alphas_h
  beta_samps[iter,] <- cur_betas_h
  theta_samps[,,iter] <- cur_thetas_h
}

library(coda)
plot(as.mcmc(alpha_samps[,1]))
plot(as.mcmc(alpha_samps[,2]))
plot(as.mcmc(beta_samps[,1]))
plot(as.mcmc(beta_samps[,2]))
#pick a random theta_ij
plot(as.mcmc(theta_samps[2,2,]))

#completely negative interval -- indicates model 2 is better -- correct
quantile(colMeans(theta_samps[1,,10001:N_mcmc]-theta_samps[2,,10001:N_mcmc]), .025)
quantile(colMeans(theta_samps[1,,10001:N_mcmc]-theta_samps[2,,10001:N_mcmc]), .975)
```